Robust positively invariant sets for state dependent and scaled disturbances

### 2015-08-04 postreview.tex
checked and reduced to 6 pages

b\; \forall\,a\in\mathcal A,\, b\in\mathcal B\}$. We use ${\bf{1}}$ to denote the column vector of ones 
in appropriate dimensions, and $\wedge$ is the logical AND operator, index sets are denoted by $\mathcal I

@@

  (1-\lambda)\underbrace{(\{z_2\}\oplus T(z_2))}_{\subseteq S}\\
  \subseteq& Z

@@

Notice that $v_{0,i}^\ast(x)$ is not necessarily given by a unique maximiser, 
but is the solution of a multi-parametric linear program and hence will be given 

@@

Since each vertex is a piece-wise affine function of $x$, the maximum $v_{0,i}^\ast(x)$ will also
be piece-wise affine
and therefore the set $\mathcal X_1=\mathcal X_0 \cap \{x:\genmat_{0,i}\Psi x + v_{0,i}^\ast(x) \leq 

@@

has the representation $\mathcal X_1 = \{x:\genmat_{1,i}x\leq\genvec_{1,i}\,\forall i\in\mathcal I_1\}$.

@@

	v_{l,i}^\ast(x)&=\max_j \ \genmat_{0,i}\Psi^{l-1}v_j(x)
	 = \begin{array}[t]{rl} \displaystyle\max_v & \genmat_{0,i}\Psi^{l-1}v \\ \text{s.t.}& v\in\mathcal V(x)
   \end{array} \\ 
   &= \begin{array}[t]{rl} \displaystyle\max_{\tilde v} & \genmat_{0,i}\tilde v\\ \text{s.t.}& \tilde v\in 

@@

@@

$\mathcal X_k$ is contained in a ball of finite radius, and hence we conclude from~\eqref{eq:set:iteration:state:dependent:constraints} that 
$\mathcal X_{k+1} = \mathcal X _k$ for all $k\geq N$ and therefore $\mathcal X _N = \mathcal X^\infty$ for finite $N$.

@@

\end{figure}

@@

In this section we discuss the calculation of the MRPI set for a linearised and simplified model of the magnetic 
levitation system depict in figure~\ref{fig:levitating:ball}. The system dynamics for the ball are given

@@

For illustration purposes we neglect inductive dynamics and use the current $u=i$ as an input and the position
$y$ and its first derivative $\dot y$ as the states, i.e. $x = (y,\dot y)^T$. We find that any equilibrium has 

@@

the Euler formula $x^+=x+T_s f(x,u) =:\tilde f(x,u)$ giving

@@

\Delta x^+ = A \Delta x + B \Delta u , \quad
\Biggl\{\begin{aligned} A &= I+T_s\frac{\partial f}{\partial  x}(\hat
  x,\hat u) \\
B &= T_s \frac{\partial f}{\partial u}(\hat x,\hat u)
\end{aligned}

@@

@@

Mean Value Theorem for vector-valued functions (e.g.~\cite{Apostol:1974}).

@@

$\tilde{x} = x - \hat{x}$, $\tilde{u} = u - \hat{u}$ we obtain the successor state:

@@

x^+ &= \tilde{f}(\hat{x} + \tilde{x},\hat{u} + \tilde{u}) 
&= \tilde f(\hat x, \hat u) 

@@

&\quad + \int_0^1\frac{\partial\tilde f}{\partial u}(\hat x + t\tilde x,\hat u+
t\tilde{u}) dt\cdot \tilde{u} \\
\Leftrightarrow 
\tilde x^+ &= A\tilde{x} + B \tilde{u} +\Bigl(

@@

&\quad +\Bigl(\int_0^1\frac{\partial\tilde f}{\partial u}(\hat x +

@@

@@

Suppose that for all $x\in\mathcal X$ and $u\in\mathcal U$, where
$\mathcal X$ and $\mathcal U$ are compact sets, $f(x,u)$ is
continuously differentiable. Then

@@

Clearly, we can then introduce the element\-wise disturbance bound 

@@

With this set we can
guarantee that $\tilde x^+ = A\tilde x + B\tilde u + v$ accounts for all nonlinearities within $\mathcal X
\times\mathcal U$ if we constrain $v\in\mathcal V(\tilde x,\tilde u)$. For general nonlinear systems
finding the extremal values of $(H^x,H^u)$ cannot be done easily.

@@

levitating ball: $T_s=30ms, C=1, m=100g, \hat x_1 = 50mm$ and $\mathcal X=\{x:
\abs{x_1- \hat x_1}\leq 1mm\wedge \abs{x_2}\leq 105\frac{mm}{s}\}$, $\mathcal U=\{u:\abs{ u-\hat u}\leq10mA\}$.

@@

nonlinearities as we will see in Section~\ref{sec:second:example}.

@@

\caption{The maximal robust positively invariant set for the levitating ball system.}

@@

@@

\small
\begin{array}{rlcrlcrl}

@@

We then use the representation~\eqref{eq:closed:form:parametrised:iterates} to argue that $D_k$ as
in~\eqref{eq:parametrised:Dk} grows exponentially and therefore contains any compact set after a finite number

@@

$\frac{1}{1-\mathcal G_i\hat\theta}$; clearly we can re-scale the rows to accommodate any other choice of $\hat\theta>-1$.

@@

First we present the parametrised MRPI
set for the levitating ball. To obtain results comparable to those in Section~\ref{sec:first:example} we derive fixed bounds
on the effect of nonlinearities on each state by finding a fixed set containing the set $\mathcal V(x,u)$ 
in~\eqref{eq:definition:element:wise:constraints:on:nonlinearities} for all $(x,u)\in\mathcal X\times\mathcal U$. 
This yields a constant set $\mathcal V$ which is non-symmetric around the origin due to nonlinearity. We also introduce
a scaling parameter $\alpha$ such that $\mathcal U(0)=\mathcal U$ for $\alpha=0$. For this setup we obtain the MRPI
set $\mathcal Z^\infty=\{(x,\theta,\alpha): \Lambda_i^x x + \Lambda_i^\theta \theta + \Lambda_i^\alpha \alpha\leq
\lambda_i\,\forall i\leq m_\infty\}$. Notice that since $\alpha$ was introduced as a scaling parameter
for $\mathcal U(\alpha)=\{u: Fu\leq (1+\alpha){\bf{1}}\}$ all entries in $\mathcal H$ are non positive and
remain unchanged throughout the computation of $\mathcal Z^\infty$ so that $\Lambda^\alpha$ has only non
positive entries, i.e. increasing $\alpha$ will enlarge the parametrised MRPI set $\mathcal Z^\infty\vert_{\alpha_1}
\subseteq\mathcal Z^\infty\vert_{\alpha_2}$ for $\alpha_1\leq\alpha_2$.

First we compare the MRPI set $\mathcal X^\infty$ obtained with state dependent disturbance 
constraints with the parametrised set $\mathcal Z^\infty$. To do this we compute the scaling parameter 
$\alpha_{\min}$ such that $\mathcal X^\infty\subseteq\mathcal Z^\infty\vert_{(\theta=0,\alpha=\alpha_{\min})}$
by solving $m_\infty$ linear programs
%
%
\[
	\gamma_i^\ast = \begin{aligned}[t]
	\max_x & \ \Lambda_i^x x\\
	\text{s.t.}& \ x\in\mathcal X^\infty
	\end{aligned}
\]
%
%
The minimal value for $\alpha_{\min}$ is then given by the maximal $\alpha$ satisfying $\gamma_i^\ast + 
\Lambda_i^\theta\cdot 0 + \Lambda_i^\alpha \alpha \leq \lambda_i$. Solving this we obtain the 
minimal value $\alpha_{\min} = 1.7555$ and the MRPI sets shown in Figure~\ref{fig:minimal:scaling:comparison:MRPIs}.
The computation terminates after seven iterations and produces a polyhedron $\mathcal Z^\infty$ supported by $m_\infty=10$
planes.
Caution is advised when interpreting $\alpha>0$, since the system is nonlinear.
extrapolations reveal little insight, however $\alpha_{\min} > 0$ indicates that the current set of inputs $\mathcal U(0)=\mathcal U$ 
is not large enough to cope with the perturbation set given by upper bounds of the nonlinear effects.
%
%
\begin{figure}
\begin{lpic}{minimalAlphaForConstantDisturbance(.65,)}
{\tiny
\lbl[r]{7,87;$1.5$}
\lbl[r]{7,69;$1$}
\lbl[r]{7,51;$0.5$}
\lbl[r]{7,34;$0$}
\lbl[r]{7,16;$-0.5$}
\lbl[t]{15,2;$-0.06$}
\lbl[t]{27,2;$-0.05$}
\lbl[t]{39,2;$-0.04$}
\lbl[t]{51,2;$-0.03$}
\lbl[t]{63,2;$-0.02$}
\lbl[t]{75,2;$-0.01$}
\lbl[t]{89,2;$0$}
\lbl[t]{100,2;$0.01$}
\lbl[t]{113,2;$0.02$}
}
{\footnotesize
\lbl[r]{5,45,90;$\tilde x_2$}
\lbl[t]{121,3;$\tilde x_1$}
}
\lbl[l]{93,32;$\longleftarrow\mathcal X^\infty$}
\lbl[r]{68,43;$\mathcal Z^\infty\vert_{(0,\alpha_{\min})}\longrightarrow$}
\end{lpic}
\caption{The minimal scaling of the constant disturbance set such that the associated MRPI set contains
the MRPI set for piecewise affine disturbances is $\alpha_{\min}=1.7555$. Note that since the nonlinearities of the system
 dynamics are not symmetrical around the equilibrium the MRPI set with fixed constraints on the disturbance
 is non-symmetric.}
 \label{fig:minimal:scaling:comparison:MRPIs}
\end{figure}

@@

$x^TPx - ((A+BK)x+Dw)^TP((A+BK)x+Dw)\geq x^T(Q+K^TRK)x -\gamma^2 w^Tw$, i.e. $K_w = (\gamma^2-D^TPD)D^TP(A+BK)$.
We use $Q = \text{diag}(2,1)$ and $R = I$. Using these numerical values we determine the MRPI set 

@@

i.e. the third input constraint.
Similarly to the first example we might also want to know how much disturbance the closed loop system can tolerate 

@@

\end{figure}

@@

In this paper we discussed extensions to existing computational methods to determine MRPI sets for linear systems subject 

>>>>

b\; \forall\,a\in\mathcal A,\, b\in\mathcal B\}$. We denote the column vector of ones 
in appropriate dimensions as ${\bf{1}}$, the logical AND operator as $\wedge$, and index sets as  $\mathcal I

@@

  (1-\lambda)\underbrace{(\{z_2\}\oplus T(z_2))}_{\subseteq S}
  \subseteq Z

@@

Here $v_{0,i}^\ast(x)$ is not necessarily given by a unique maximiser, 
but is the solution of a multi-parametric linear program and hence is given 

@@

Since each vertex is a piece-wise affine function of $x$, the maximum $v_{0,i}^\ast(x)$ is also piece-wise affine
so the set $\mathcal X_1=\mathcal X_0 \cap \{x:\genmat_{0,i}\Psi x + v_{0,i}^\ast(x) \leq 

@@

has representation $\mathcal X_1 = \{x:\genmat_{1,i}x\leq\genvec_{1,i}\,\forall i\in\mathcal I_1\}$.

@@

	v_{l,i}^\ast(x)=\max_j \ \genmat_{0,i}\Psi^{l-1}v_j(x)
   = \begin{array}[t]{rl} \displaystyle\max_{\tilde v} & \genmat_{0,i}\tilde v\\ \text{s.t.}& \tilde v\in 

@@

% \begin{align*}
% 	v_{l,i}^\ast(x)&=\max_j \ \genmat_{0,i}\Psi^{l-1}v_j(x)
% 	 = \begin{array}[t]{rl} \displaystyle\max_v & \genmat_{0,i}\Psi^{l-1}v \\ \text{s.t.}& v\in\mathcal V(x)
%    \end{array} \\ 
%    &= \begin{array}[t]{rl} \displaystyle\max_{\tilde v} & \genmat_{0,i}\tilde v\\ \text{s.t.}& \tilde v\in 
%    \Psi^{l-1}\mathcal V(x) \end{array}
% \end{align*}

@@

$\mathcal X_k$ is contained in a ball of finite radius, and we conclude from~\eqref{eq:set:iteration:state:dependent:constraints} that 
$\mathcal X_{k+1} = \mathcal X _k$ for all $k\geq N$ and hence $\mathcal X _N = \mathcal X^\infty$ for finite $N$.

@@

\vspace{-2mm}\end{figure}

@@

This section discusses the calculation of the MRPI set for a linearised, simplified model of the magnetic 
levitation system depicted in figure~\ref{fig:levitating:ball}. The system dynamics for the ball are given

@@

For illustration purposes we neglect inductive dynamics, using the current $u=i$ as the control input and the position
$y$ and its first derivative $\dot y$ as the states: $x = (y,\dot y)^T$. We find that any equilibrium has 

@@

the Euler formula $x^+=x+T_s f(x,u) =:\tilde f(x,u)$ giving $\Delta x^+ = A \Delta x + B \Delta u$,

@@

A = I+T_s\frac{\partial f}{\partial  x}(\hat
  x,\hat u) , \quad
B = T_s \frac{\partial f}{\partial u}(\hat x,\hat u)

@@

% \[
% \Delta x^+ = A \Delta x + B \Delta u , \quad
% \Biggl\{\begin{aligned} A &= I+T_s\frac{\partial f}{\partial  x}(\hat
%   x,\hat u) \\
% B &= T_s \frac{\partial f}{\partial u}(\hat x,\hat u)
% \end{aligned}
% \]

@@

Mean Value Theorem 
(e.g.~\cite{Apostol:1974}).

@@

$\tilde{x} = x - \hat{x}$, $\tilde{u} = u - \hat{u}$, we obtain the successor state $x^+ = \tilde{f}(\hat{x} + \tilde{x},\hat{u} + \tilde{u})$ as

@@

\begin{multline*}
%x^+ &= \tilde{f}(\hat{x} + \tilde{x},\hat{u} + \tilde{u}) 
%\\
x^+= \tilde f(\hat x, \hat u) 

@@

+ \int_0^1\frac{\partial\tilde f}{\partial u}(\hat x + t\tilde x,\hat u+
t\tilde{u}) dt\cdot \tilde{u}
\end{multline*}
and hence
\begin{multline*}
\tilde x^+ = A\tilde{x} + B \tilde{u} +\Bigl(

@@

+\Bigl(\int_0^1\frac{\partial\tilde f}{\partial u}(\hat x +

@@

\end{multline*}
% \begin{align*}
% %x^+ &= \tilde{f}(\hat{x} + \tilde{x},\hat{u} + \tilde{u}) 
% %\\
% x^+&= \tilde f(\hat x, \hat u) 
% + \int_0^1\frac{\partial\tilde f}{\partial x}(\hat x + t\tilde x,
% \hat u + t\tilde u) dt \cdot \tilde x  \\
% &\quad + \int_0^1\frac{\partial\tilde f}{\partial u}(\hat x + t\tilde x,\hat u+
% t\tilde{u}) dt\cdot \tilde{u} \\
% \Leftrightarrow 
% \tilde x^+ &= A\tilde{x} + B \tilde{u} +\Bigl(
% \int_0^1\frac{\partial\tilde f}{\partial x}(\hat x + t\tilde x,\hat u
%              + t\tilde{u})dt - A \Bigr)\tilde x 
% \\ 
% &\quad +\Bigl(\int_0^1\frac{\partial\tilde f}{\partial u}(\hat x +
%   t\tilde x,\hat u + t\tilde{u})dt - B\Bigr)\tilde{u}
% \end{align*}

@@

If $f(x,u)$ is
continuously differentiable for all $x\in\mathcal X$ and $u\in\mathcal U$, where
$\mathcal X$, $\mathcal U$ are compact sets, then

@@

Thus we can introduce the element\-wise disturbance bound 

@@

Therefore $\tilde x^+ = A\tilde x + B\tilde u + v$ accounts for all nonlinearities within $\mathcal X
\times\mathcal U$ if $v\in\mathcal V(\tilde x,\tilde u)$. For general nonlinear systems
finding the extremal values of $(H^x,H^u)$ is non-trivial.

@@

levitating ball: $T_s=30\,\text{ms}$, $C=1$, $m=100\,\text{g}$, $\hat x_1 = 50\,\text{mm}$ and $\mathcal X=\{x:
\abs{x_1- \hat x_1}\leq 1\,\text{mm}\wedge \abs{x_2}\leq 105\,\text{mm}/\text{s}\}$, $\mathcal U=\{u:\abs{ u-\hat u}\leq10\,\text{mA}\}$.

@@

nonlinearities (this will be shown in Section~\ref{sec:second:example}).

@@

\caption{The maximal robust positively invariant set for the levitating ball.}

@@

% \[
% \begin{array}[t]{@{}rl@{}}
% \max_v&\mathcal F_i \Psi^l v \\
% \text{s.t.}&v\in\bar{\mathcal{V}}
% \end{array}
% =
% \begin{array}[t]{@{}rl@{}}
% \max_{\tilde v}&\mathcal F_i \tilde  v \\
% \text{s.t.}&\Psi^{-l} \tilde v\in\bar{\mathcal{V}}
% \end{array}
% =
% \begin{array}[t]{@{}rl@{}}
% \max_{\tilde v}&\mathcal F_i \tilde v \\ 
% \text{s.t.}& \tilde v\in\Psi^l\bar{\mathcal{V}}
% \end{array}
% \]

@@

\begin{array}{@{}rl@{\hspace{0.7em}}c@{\hspace{0.7em}}rl@{\hspace{0.7em}}c@{\hspace{0.7em}}rl@{}}

@@

We then use the representation~\eqref{eq:closed:form:parametrised:iterates} to argue that $D_k$ 
in~\eqref{eq:parametrised:Dk} grows exponentially and hence contains any compact set after a finite number

@@

$(1-\mathcal G_i\hat\theta)^{-1}$; clearly we can re-scale the rows to accommodate any other choice of $\hat\theta>-1$.

@@

First we present the parametrised MRPI
set for the levitating ball. To obtain results comparable to those in Section~\ref{sec:first:example} we derive fixed bounds
on the effect of nonlinearities on each state by finding a set $\mathcal V$ that contains $\mathcal V(x,u)$ 
in~\eqref{eq:definition:element:wise:constraints:on:nonlinearities} for all $(x,u)\in\mathcal X\times\mathcal U$. 
%This yields a fixed set $\mathcal V$ which is non-symmetric around the origin due to nonlinearity.
We also introduce
a scaling parameter $\alpha$ such that $\mathcal U(0)=\mathcal U$ for $\alpha=0$. The MRPI set then has the form $\mathcal Z^\infty=\{(x,\theta,\alpha): \Lambda_i^x x + \Lambda_i^\theta \theta + \Lambda_i^\alpha \alpha\leq
\lambda_i\,\forall i\leq m_\infty\}$. Since $\alpha$ was introduced as a scaling parameter
for $\mathcal U(\alpha)=\{u: Fu\leq (1+\alpha){\bf{1}}\}$, all entries in $\mathcal H$ are non positive and
remain unchanged throughout the computation of $\mathcal Z^\infty$; hence $\Lambda^\alpha$ has only non
positive entries, and increasing $\alpha$ enlarges the parametrised MRPI set $\mathcal Z^\infty\vert_{\alpha_1}
\subseteq\mathcal Z^\infty\vert_{\alpha_2}$ for $\alpha_1\leq\alpha_2$.

To compare the MRPI set $\mathcal X^\infty$ for state dependent disturbance 
constraints with the parametrised set $\mathcal Z^\infty$, we determine
$\alpha_{\min}$ such that $\mathcal X^\infty\subseteq\mathcal Z^\infty\vert_{(\theta=0,\alpha=\alpha_{\min})}$
by solving $m_\infty$ linear programs:
%
%
\[
	% \gamma_i^\ast = 
	% \max_x \ \Lambda_i^x x \ \ 
	% \text{s.t.} \ x\in\mathcal X^\infty
	\gamma_i^\ast = \begin{aligned}[t]
	\max_x & \ \Lambda_i^x x\\
	\text{s.t.}& \ x\in\mathcal X^\infty
	\end{aligned}
\]
%
%
The minimal value for $\alpha_{\min}$ is then given by the maximal $\alpha$ satisfying $\gamma_i^\ast + 
\Lambda_i^\theta\cdot 0 + \Lambda_i^\alpha \alpha \leq \lambda_i$. This gives the 
minimal value of $\alpha_{\min} = 1.7555$ and the MRPI sets shown in Figure~\ref{fig:minimal:scaling:comparison:MRPIs}.
The computation terminates after seven iterations and produces a polyhedron $\mathcal Z^\infty$ supported by $m_\infty=10$
planes.
Caution is advised when interpreting $\alpha>0$, since the system is nonlinear,
%extrapolations reveal little insight, however 
however $\alpha_{\min} > 0$ indicates that the input constraint set $\mathcal U(0)=\mathcal U$ 
is too small to cope with a fixed disturbance set bounding the linearisation errors.
%
%
\begin{figure}
\begin{lpic}{minimalAlphaForConstantDisturbance(.65,)}
{\tiny
\lbl[r]{7,87;$1.5$}
\lbl[r]{7,69;$1$}
\lbl[r]{7,51;$0.5$}
\lbl[r]{7,34;$0$}
\lbl[r]{7,16;$-0.5$}
\lbl[t]{15,2;$-0.06$}
\lbl[t]{27,2;$-0.05$}
\lbl[t]{39,2;$-0.04$}
\lbl[t]{51,2;$-0.03$}
\lbl[t]{63,2;$-0.02$}
\lbl[t]{75,2;$-0.01$}
\lbl[t]{89,2;$0$}
\lbl[t]{100,2;$0.01$}
\lbl[t]{113,2;$0.02$}
}
{\footnotesize
\lbl[r]{5,45,90;$\tilde x_2$}
\lbl[t]{121,3;$\tilde x_1$}
}
\lbl[l]{93,32;$\longleftarrow\mathcal X^\infty$}
\lbl[r]{68,43;$\mathcal Z^\infty\vert_{(0,\alpha_{\min})}\longrightarrow$}
\end{lpic}
\caption{MRPI sets for the minimally scaled fixed disturbance set and for the piecewise affine disturbance set. The MRPI set for fixed disturbance bounds is non-symmetric because the nonlinearities of the system dynamics are not symmetrical about the equilibrium.}
%The minimal scaling of the constant disturbance set such that the associated MRPI set contains %the MRPI set for piecewise affine disturbances is $\alpha_{\min}=1.7555$. 
%Note that since the nonlinearities of the system dynamics are not symmetrical around the equilibrium the MRPI set with fixed constraints on the disturbance
% is non-symmetric.}
\label{fig:minimal:scaling:comparison:MRPIs}
\vspace{-2mm}\end{figure}

@@

$x^TPx - ((A+BK)x+Dw)^TP((A+BK)x+Dw)\geq x^T(Q+K^TRK)x -\gamma^2 w^Tw$, i.e. $K_w = (\gamma^2-D^TPD)D^TP(A+BK)$
for $Q = \text{diag}(2,1)$ and $R = I$. Using these numerical values we determine the MRPI set 

@@

i.e.~the third input constraint.
As for the first example we might also want to know how much disturbance the closed loop system can tolerate 

@@

\vspace{-4mm}\end{figure}

@@

This paper discusses extensions to existing computational methods for determining MRPI sets for linear systems subject 
